### 2020-08-21 msfp2020/msfp2020.tex
A start to the slides for MSFP 2020

\documentclass{beamer}

\usepackage{fontspec}
\usepackage{minted}
\usepackage{xspace}

\setmainfont{TeX Gyre Heros}
\setmonofont{Noto Sans Mono}
\usetheme{Madrid}

\let \parskip=\baselineskip
\setminted{fontsize=\footnotesize}

\newcommand{\Agda}[1]{\texttt{\footnotesize #1}\xspace}
\newcommand{\Free}{\Agda{Free}}
\newcommand{\match}{\Agda{match}}
\newcommand{\Nondet}{\Agda{Nondet}}
\newcommand{\Regex}{\Agda{Regex}}
\newcommand{\Set}{\Agda{Set}}

\title[Combining predicate transformer semantics]{Combining predicate transformer semantics for effects: a case study in parsing regular languages}
\author{Anne Baanen \and Wouter Swierstra}
\date{September 1, 2020}

\begin{document}

\titlepage

\begin{frame}{The history}

\emph{Predicate transformer semantics} for imperative programs have been around for 45 years.
A statement becomes a map from postcondition (a predicate on the state space) to the weakest precondition that ensures the postcondition will be satisfied.

\pause

\emph{Parser combinators} have been around in functional programming for 35 years.
Parsing becomes an effectful (monadic) operation, nondeterministically modifying state (the unparsed part of the string).

\pause

\emph{Algebraic effects} have recently gained traction in the programming language community.

\end{frame}

\begin{frame}[fragile]{Algebraic effects}

Algebraic effects separate the syntax and semantics of effects.

\begin{itemize}
	\item The syntax describes the sequencing of the primitive operations
	\item The semantics assigns meaning to these operations
\end{itemize}

\begin{minted}{Agda}
data Free (C : Set) (R : C -> Set) : Set -> Set where
	Pure : a -> Free C R a
	Op : (c : C) -> (k : R c -> Free C R a) -> Free C R a

data CNondet where
	Fail : CNondet
	Choice : CNondet
RNondet : CNondet -> Set
RNondet Fail = ⊥
RNondet Choice = Bool

Nondet = Free CNondet RNondet
\end{minted}
\end{frame}

\begin{frame}[fragile]{Semantics for algebraic effects}
\emph{Handlers} give semantics for the \Free monad naturally as a fold:

\begin{minted}{Agda}
handleList : Nondet a -> List a
handleList (Pure x) = [x]
handleList (Op Fail k) = []
handleList (Op Choice k) = k True ++ k False
\end{minted}

The generic fold that computes a predicate of type \Set:

\begin{minted}{Agda}
[[_]] : Free C R a -> ((c : C) -> (R c -> Set))
      -> (a -> Set) -> Set
[[ Pure x ]] alg P = P x
[[ Op c k ]] alg P = alg c (λ x -> [[ k x ]] alg P)
\end{minted}
\end{frame}

\begin{frame}[fragile]{Predicate transformer semantics}

A predicate transformer for commands \Agda{C} and responses \Agda{R}
is a function from postconditions of type \Agda{R -> Set} to preconditions of type \Agda{C -> Set}.
If \Agda{R} depends on \Agda{C}, this becomes:

\begin{minted}{Agda}
pt C R = (c : C) -> (R c -> Set) -> Set
\end{minted}

The type of the algebra passed to \Agda{[[\_]]} is exactly \Agda{pt C R}.
We have assigned \emph{predicate transformer semantics} to algebraic effects.

\end{frame}

\begin{frame}[fragile]{Predicate transformer semantics for \Nondet}

For nondeterminism, there are two canonical choices of predicate transformer semantics.

\Agda{ptAll} requires that all potential results satisfy the postcondition:
\begin{minted}{Agda}
ptAll Fail k = ⊤
ptAll Choice k = k True ∧ k False
\end{minted}

\Agda{ptAny} requires that there is at least one outcome that satisfies the postcondition:
\begin{minted}{Agda}
ptAny Fail k = ⊥
ptAny Choice k = k True ∨ k False
\end{minted}

\end{frame}

\begin{frame}[fragile]{Parsing regular expressions}

Consider the following syntax of regular expressions and their parse trees:
\begin{minted}{Agda}
data Regex : Set where
    Empty : Regex
    Epsilon : Regex
    Singleton : Char → Regex
    _ | _ : Regex → Regex → Regex
    _ · _ : Regex → Regex → Regex
    _ * : Regex → Regex

Tree : Regex -> Set
Tree Empty = ⊥
Tree Epsilon = ⊤
Tree (Singleton _) = Char
Tree (l | r) = Either (Tree l) (Tree r)
Tree (l · r) = Pair (Tree l) (Tree r)
Tree (r *) = List (Tree r)
\end{minted}

Let's write a parser returning these \Agda{Tree}s.
\end{frame}

\begin{frame}[fragile]{Parsing regular expressions}
A nondeterministic \Regex matcher is straightforward to write:
\begin{minted}{Agda}
match : (r : Regex) -> String -> Nondet (Tree r)
match Empty xs = Op Fail λ()
match Epsilon Nil = Pure tt
match Epsilon (_ :: _) = Op Fail λ()
match (Singleton c) xs =
    if xs = [c] then Pure c else Op Fail λ()
match (l | r) xs = Op Choice (λ b,
    if b then Inl <$> match l xs else Inr <$> match r xs)
match (l · r) xs = do
    (ys, zs) <- allSplits xs
    (,) <$> match l ys <*> match r zs
\end{minted}

\pause
... until we reach the last case:
\begin{minted}{Agda}
match (r *) xs = match (r · (r *)) xs
    -- Error: does not terminate
\end{minted}

\end{frame}

\begin{frame}[fragile]{Parsing regular expressions}
For now, we will write:
\vspace{-\baselineskip}
\begin{minted}{Agda}
match (r *) xs = Op Fail λ()
\end{minted}
\vspace{\baselineskip}

\pause

To verify our implementation, we take a specification consisting of precondition and postcondition:
\begin{minted}{Agda}
pre : Regex -> String -> Set
pre r xs = hasNo* r

post : (r : Regex) -> String -> Tree r -> Set
post r xs t = Match r xs t
\end{minted}

And check that \Agda{match} \emph{refines} this specification.

\end{frame}

\begin{frame}[fragile]{Refinement calculus}

A predicate transformer \Agda{pt1} \emph{is refined by} \Agda{pt2} if \Agda{pt₂} satisfies more postconditions than \Agda{pt₁}:
\vspace{-\baselineskip}
\begin{minted}{Agda}
_⊑_ : (pt1 pt2 : (a -> Set) -> Set) -> Set
pt1 ⊑ pt2 = ∀ P -> pt1 P -> pt2 P
\end{minted}

The relation \Agda{S ⊑ T} expresses that the program \Agda{T} is ``better'' than \Agda{S}:
\Agda{S} can be replaced with \Agda{T} everywhere, and all postconditions will still hold.

\vspace{\baselineskip}

By assigning a predicate transformer to specifications, we can also relate specifications and programs:
\vspace{-\baselineskip}
\begin{minted}{Agda}
[[_,_]] : (pre : Set) (post : a -> Set) -> (a -> Set) -> Set
[[ pre , post ]] P = pre ∧ ∀ x, post x -> P x
\end{minted}

This is the `weakest precondition' necessary so that the desired post- condition \Agda{P} holds:
\Agda{pre} should hold and any result satisfying \Agda{post} should imply the postcondition \Agda{P}.

\end{frame}

\begin{frame}[fragile]{Verification}

With these ingredients, the correctness statement of \match becomes:
\begin{minted}{Agda}
matchSound : (r : Regex) (xs : String) ->
    [[ pre r xs , post r xs ]] ⊑ [[ match r xs ]] ptAll
\end{minted}

The proof proceeds by case distinction and is uncomplicated, until we need to reason about the monadic bind operator \Agda{\_>>=\_}.

\pause

The missing ingredient is the rule of consequence:
\begin{minted}{Agda}
consequence : ∀ pt (S : Free es a) (f : a → Free es b) →
    [[ S ]] pt (λ x → [[ f x ]] pt P) ≡ [[ mx >>= f ]] pt P
\end{minted}

\end{frame}

\begin{frame}[fragile]{Adding effects}
The problem with \match is that implementing the Kleene star requires the effect of \emph{general recursion}.

We can add more effects to the free monad by choosing the command and response types from a list of \emph{effect signatures}:
\begin{minted}{Agda}
data Free (es : List Sig) : Set -> Set where
	Pure : a -> Free es a
	Op : (i : mkSig C R ∈ es) (c : C) -> (k : R c -> Free C R a) -> Free C R a
\end{minted}

We will add two new effects along with nondeterminism: general recursion and parsing.
\end{frame}

\begin{frame}[fragile]{Adding effects}
The \Agda{Rec I O} effect represents a recursive function of type \Agda{(i : I) -> O i} calling itself.
The commands are the arguments to the function and the responses are the returned values.
\begin{minted}{Agda}
Rec : (I : Set) (O : I -> Set) -> Sig
Rec I O = mkSig I O
\end{minted}
To specify the semantics of \Agda{Rec}, we need an invariant of type \Agda{(i : I) -> O i -> Set},
specifying which values of type \Agda{O i} can be returned from a call with argument \Agda{i : I}.
\begin{minted}{Agda}
ptRec inv i P = ∀ o -> inv i o -> P o
\end{minted}
\end{frame}

\begin{frame}[fragile]{Adding effects}
The \Agda{Parser} effect represents a stateful parser with one command:
advance the input string by one character.

\begin{minted}{Agda}
Parser : Sig
Parser = mkSig ⊤ (λ _ -> Maybe Char)
\end{minted}

The semantics of \Agda{Parser} are stateful: returning the next character of the input string requires keeping track of the remaining characters.
This state can be found in the \Agda{ptParser} semantics as an extra argument to the predicates.
\begin{minted}{Agda}
ptParser : ⊤ -> (Maybe Char -> String -> Set) -> String -> Set
ptParser _ P Nil = P Nothing Nil
ptParser _ P (x :: xs) = P (Just x) xs
\end{minted}

\end{frame}

\begin{frame}[fragile]{Defining a derivative-based matcher}
\begin{minted}{Agda}
dmatch : (Forall(es)) ⦃ iP : Parser ∈ es ⦄ ⦃ iND : Nondet ∈ es ⦄  -> (RecArr Regex es tree)
dmatch r = symbol >>= maybe
    (λ x -> integralTree r <$> call (hiddenInstance(∈Head)) (d r /d x))
    (if p <- matchEpsilon r then Pure (Sigma.fst p) else (hiddenConst(fail)))
\end{minted}
\end{frame}

\begin{frame}[fragile]{Termination checking}
\begin{minted}{Agda}
terminates-in : (Forall(I O es a)) (pts : PTs es) (f : (RecArr I es O)) (S : Free (Rec I O :: es) a) → Nat → Set
terminates-in pts f (Pure x)            n         = ⊤
terminates-in pts f (Op ∈Head c k)      Zero      = ⊥
terminates-in pts f (Op ∈Head c k)      (Succ n)  = terminates-in pts f (f c >>= k) n
terminates-in pts f (Op (∈Tail i) c k)   n        =
    lookupPT pts i c (λ x → terminates-in pts f (k x) n)
\end{minted}
\end{frame}

\begin{frame}[fragile]{Refining \match with \Agda{dmatch}}
\begin{minted}{Agda}
dmatchSound : ∀ r xs -> (wpMatch (match (hiddenInstance(∈Head)) (r , xs))) ⊑ (wpMatch (dmatch' (hiddenInstance(∈Head)) (r , xs)))
\end{minted}
\end{frame}

\end{document}